---
title: "Gaussian Processes"
author: "Michael Budjan, Moritz Haas, Konstantin Klumpp, Tim Reitze"
date: "`r Sys.Date()`"
header-includes:
   - \usepackage{bbm}
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(gaussianprocess)
```
> "You shouldn't feel ashamed about your code - if it solves the problem, it's perfect just the way it is. But also, it could always be better." — @hadleywickham
 [via](https://twitter.com/allimoberger/status/1085268564821585921)
 
A Gaussian Process is a stochastic process $(X_t)_{t \in T}$ with $\mathbb{E}\left[ X_t \right] = 0$ and for all finite subsets $S \subset T$, $(X_t)_{t \in S}$ has a multivariate normal distribution.

## Regression
```{r, eval=FALSE}
Gaussian <- GPR$new(X, y, cov_func, noise)
```
Some covariance functions have already been implemented, they can be used by selecting the corresponding subclass.

```{r, eval=FALSE}
Gaussian <- GPR.sqrexp$new(X, y, l = 1, noise = 0.5)
```

```{r, eval=FALSE}
Gaussian$plot(seq(-10, 10, by = 0.05))
```

```{r dpi=1000, out.width="100%", echo=FALSE, fig.width = 5, fig.asp = .62}
X <- matrix(seq(-5,5,by = 0.5), nrow = 1)
noise <- 0.5
y <- c(0.1*X^3 + rnorm(length(X), 0, noise))
Gaussian <- GPR.sqrexp$new(X, y, 1, noise)
Gaussian$plot(seq(-10,10, by = 0.05))
```

## Classification


